---
knit: bookdown::preview_chapter
---

# Description of Simulation Framework


Here, I simulate some data in which the residuals are correlated. 

Thus $\beta \sim N(0,Uk)$
and $\hat{\beta} \sim N(0,Uk+Vj)$

Such that $\hat{\beta} = \beta + E$ where $E \sim N(0,Vj)$

```{r, eval=F,echo=TRUE}
rm(list=ls())

library('mash')

sim.with.error=function(J,d=44,betasd=1,esd=0.11,n=400,rho=0.8){
   n=n
  covmat=readRDS(system.file('simdata/covmatforsimulation.rds', package = 'mash'))[2:9]
  covmat=lapply(seq(1:length(covmat)),function(x){covmat[[x]]/max(diag(covmat[[x]]))})
  
  
  
  library("mvtnorm")
  library("MASS")
  K=length(covmat)
  
  
  if(n!=0){
  z = sample(K,n,replace=TRUE)
  omega=abs(rnorm(n,mean=0,sd=betasd))##effect size variance can be big or small
  beta=t(sapply(seq(1:n),function(j){
    k=z[j]
    o=omega[j]
    mvrnorm(1,mu=rep(0,d),Sigma=o*covmat[[k]])
  }))
  beta=rbind(beta,matrix(rep(0,(J-n)*d),ncol=d))}
  if(n==0){
    beta=matrix(rep(0,(J-n)*d),ncol=d)
  }
  
  s.j.r=as.matrix(abs(rnorm(d,esd,0.001)))##simulate with the same standard error for every J
  v.j.r=s.j.r%*%t(s.j.r)##now v.j.r will be the same for every J
  v.mat=rho*v.j.r+(1-rho)*diag(diag(v.j.r))#make the errors correlated
  e=rmvnorm(J,mean=rep(0,d),sigma=v.mat)
  betahat = beta + e
  s.j=matrix(rep(s.j.r),nrow(betahat),byrow=T,ncol=d)
  t.stat=betahat/abs(s.j)
  if(n!=0){
    return(list(beta=beta,betahat=betahat,component.mats=covmat,sebetahat=s.j,t.stat=t.stat,component.id=z,error=e,var.mat=v.mat,omega=omega))
  }
  if(n==0){
    return(list(beta=beta,betahat=betahat,sebetahat=s.j,t.stat=t.stat,error=e,var.mat=v.mat))
  }
}
```

You can see that in this simulation, the $V.j$ matrix for ever gene is the same for every j, and the standard errors for the tissues are all roughly equivalnet (centered aroun 0.10.)


I simulate 1000 values, 400 of which are `true` shared structured with $rho$ = 0.8 in the correlation of residual matrix.
```{r, echo=FALSE}
library('mash')
seed=4
set.seed(seed)
s=sim.with.error(J = 1000,d = 44,rho = 0.8,n=400)

attributes(s)
```



